---
title: "Exercises"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Introduction to Data Science: Data pre-processing Exercises Part 1
1. Consider the following table summarizing different studies. Classify each variable as qualitative (qual) / quantitative (quant). Indicator also if it is a discrete or a continuous variable.

2. Explain why zip codes should be considered text variables rather than numeric.

3. Consider a data set $X_1, X_2, \dots,X_{30}$ with 30 observations. Assume that each observation can be missing and assume that the probability that an observation is missing is 0.05. Moreover, missing values are spread evenly through the data and occur independently of each other. Denote by $D_i$ the random variable which takes value 1 if observation $i$ is missing and 0 otherwise. Show that the probability that the data set of 30 observations contains at least one missing value is close to 80%.

4. Consider $n_1 + m$ independent and identical random variables denoted by $X_1, X_2,\dots ,X_{n_1}$ and $X_{n_1+1},\dots,X_{n_1 + m}$. Denote by $S_1$ the random mean over the first $n_1$ random variables whereas $S_2$ denotes the random mean over all random variables:

$$S_1 = \frac{1}{n_1}\sum_{i=1}^{n_1}X_i,$$ 
$$S_2 = \frac{1}{n_1 + m}\sum_{i=1}^{n_1 + m} X_i.$$
Assume that the expectation of $X_1$ is $\mu$ and the variance is $\sigma^2$:
$$\mathbb{E}[X_1] = \mu$$
$$Var[X_1] = \sigma^2$$
(a) Show that
$$\mathbb{E}[S_1] = \mathbb{E}[S_2] = \mu$$
(b) Show that
$$\frac{Var[S_1]}{Var[S_2]} = 1 + \frac{m}{n_1}$$
(c) Assume the $X_i$ is the $i$-th observation in our data set. Moreover, assume that the last $m$ observations have to be deleted because of missing values. Give an interpretation of the results in Question a and b

5. Consider the following R code. What is the output after running this code?
```{r}
Data = data.frame(c(100,NA,90,75,110,132), c(NA,1,5,0.5,NA, "na"))
names(Data) = c("Variable 1", "Variable 2")
sum(is.na(Data))
```

6. Consider the following R code which is used to handle missing values, where Data is a vector with observations.
```{r, eval=FALSE}
sum(is.na(Data))
Data.New = replace(Data,which(is.na(Data)), mean(Data))
```

